---
title: "Spatial join with the BC CHSA"
author: "Jonathan Simkin"
date: '2022-05-28'
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

# Intro

Welcome! Here, I'll demonstrate a quick way to link the British Columbia (BC) Community Health Service Area (CHSA) shapefile, with any given point data (E.g. imagine you have health events by longitude / latidude). We'll do this through a spatial join!

I have this published on RPubs as well, you can check that out here: https://rpubs.com/jdsimkin04

For RPubs sake, here's the github repo: https://github.com/jdsimkin04/bc_chsa_spatial_join

## What is the BC CHSA shapefile?

The CHSA is the smallest health administrative area in BC. It was created to facilitate community level health analyses. The shapefile is openly available through the [BC Gov data catalogue](https://catalogue.data.gov.bc.ca/dataset/community-health-service-areas-chsa). It also contains attribute data - contextual information related to that CHSA. These include but are not limited too:

-   Corresponding health administrative areas (LHA, HSDA, HA)
-   Census population
-   Area km^2^
-   Level of urbanization

## Why would we want to link the CHSA to data?

-   To assign rural-urban status in BC to point data (e.g. health events)
-   To assign CHSA, LHA, HSDA, or HA in BC

# Spatial join rural-urban class to point data

## Download the CHSA map file

We can do this a few ways... you could go to the [BC Gov data catalogue](https://catalogue.data.gov.bc.ca/dataset/community-health-service-areas-chsa) and download the shapefile. Or you could use the download.file and unzip functions shown below.

### Libraries

install these libraries if you don't have them

```{r, results = "hide"}
library(tidyverse)
library(sf)
library(tmap)
library(tmaptools)
library(kableExtra)
library(rmapshaper)
```

### Download and unzip

Download the file from the website using the download.file and unzip function to unzip into a folder called map

```{r, eval = FALSE}
url <- "https://catalogue.data.gov.bc.ca/dataset/68f2f577-28a7-46b4-bca9-7e9770f2f357/resource/f89f99b0-ca68-41e2-afc4-63fdc0edb666/download/chsa_2018.zip"
download.file(url, "chsa_2018.zip")
unzip("chsa_2018.zip", exdir = paste0(getwd(), "/map"))
```

## Visualize

read the map files using the st_read function

```{r, results = "hide"}
chsa <-
st_read("map") %>% 
  st_make_valid() %>% 
  st_transform(., crs = 4326) %>% 
  simplify_shape(.)
```

visualize the chsa map

```{r}
tmap_mode("view")
tm_shape(chsa) +
  tm_polygons("HSDA_Name",
              id = "CHSA_Name",
              palette = "PRGn",  
              legend.show = FALSE)
```

see the data elements of the chsa map, the CHSA_UR_Cl is the rural urban classification

```{r}
chsa %>% 
  st_drop_geometry() %>% 
  str()
```

## Explore CHSA_UR_Cl

Let's look a bit closer at the CHSA_UR_Cl variable

```{r}
chsa %>% 
  st_drop_geometry() %>% 
  count(CHSA_UR_Cl) %>% 
  mutate(proportion = round(n/sum(n)*100,1))
```

```{r}
tmap_mode("view")
tm_shape(chsa) +
  tm_polygons("CHSA_UR_Cl",
              id = "CHSA_Name",
              palette = "PRGn")
```

## Sample point data

Here's a sample of point data that we want to assign CHSA rural urban class. I got these randomly from google maps.

```{r}
spatial_point_data <-
tibble(
  case_id = c(1:10),
  long = c(-122.79245404269491,
                -122.50750051771003,
                -121.47061705091572,
                -120.78489105416672,
                -119.46708958311152,
                -122.12044919737846,
                -123.9735411710232,
                -127.48038830551505,
                -122.66126844286325,
                -131.11218885196038),
  lat = c(
    49.1152475836367,  
49.117737238190344,
49.38845365756505, 
50.09960249743014, 
49.87203980873458, 
52.13143107578074, 
49.130988612031395,
50.69492877221613, 
58.80790367189857, 
57.88986797544227)
) %>% 
  st_as_sf(., coords = c("long", "lat"), crs = 4326, agr = "constant") %>% 
    st_transform(., crs = 4326)
```

## Visualize

Viz the point data

```{r}
tmap_mode("view")
tm_shape(spatial_point_data) +
  tm_dots(size = 0.1,
          col = "black")

```

Viz teh point data on the CHSAs

```{r}
tmap_mode("view")
tm_shape(chsa) +
  tm_polygons("HSDA_Name",
              id = "CHSA_Name",
              palette = "PRGn",  
              legend.show = FALSE) +
    tm_shape(spatial_point_data) +
  tm_dots(size = 0.2,
          col = "black") 

```

## Spatial join rural urban class to the point data

link cohort sf and chsa map

```{r}
int <-
  st_intersects(spatial_point_data, chsa, prepared = TRUE) #Wherever there are intersection between the two layers (the CHSA Map and the study data), we can link them! and store them in "int"

linked_data <-
spatial_point_data %>%
  mutate(
    CHSA_Name = as.character(chsa$CHSA_Name[as.numeric(as.character(int))]),
    CHSA_UR_Cl = as.character(chsa$CHSA_UR_Cl[as.numeric(as.character(int))]),
    LHA_Name = as.character(chsa$LHA_Name[as.numeric(as.character(int))]),
    HSDA_Name = as.character(chsa$HSDA_Name[as.numeric(as.character(int))]),
    HA_Name = as.character(chsa$HA_Name[as.numeric(as.character(int))])
    ) %>%
  st_drop_geometry() #drop the geometry info from the file so we can analyze the data like a norma lfile

```

# Final linked dataset

... and Voila! See the spatially joined point data with CHSA attributes.

```{r}
linked_data %>% 
  kable() %>% 
  kable_styling()
```
